---
title: "Fractribution Model - Quick start"
author: "Daniel Booth"
date: "2019-01-14"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Attribution fit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Overview

The Fractribution framework enables you to derive user-level fractional
attribution values from your marketing and event touchpoints data. This package,
**fractribution.model**, fits a customisable user-level fractional attribution
model using a simplified Shapley Value method. Additionally, and for comparison
purposes, a last-touch attribution function is also implemented in the package.

In this vignette we will explore the core functions in the package so you are
able to get started quickly.

## Examples

### Load package

First load the package. We will also load `dplyr` to use in some of the
examples:

```{r, message=FALSE}
library(fractribution.model)
library(dplyr)
```

### Paths to conversion data

The `attribution_fit()` function requires a `path_summary` input which needs to
be created from your marketing data (such as Google Analytics sessions).
Additionally, if you want to produce the attribution at the customer-level, you
need to also include a `path_customer_map` input.

You can produce this data either through your own data engineering, or through
the **fractribution.data** package (see
`?fractribution.data::run_attribution_report` for more details).

For now, for your reference, we have included examples of these two input files
in the package and we explain these below.

#### Path summary

The `path_summary` input is a dataframe with columns `path`,
`total_conversions`, and `total_non_conversions`. The last two columns are the
`path`'s aggregated conversion and non-conversion counts.

For your reference there is an `example_path_summary` dataframe loaded with the
package:

```{r}
example_path_summary
```

#### Path customer map

The `path_customer_map` input is is a dataframe mapping from `path` to
`customer_id`.

Again there is a reference, `example_path_customer_map`, loaded with the
package:

```{r}
example_path_customer_map
```

### Attribution fit

The `attribution_fit()` function runs through the counterfactual shapley value
algorithm to fit an attribution model for the paths and their conversion
probabilities.

As mentioned above, you can produce the attribution at the `customer_id` or
`path` level. To control this use the `path_level_only` argument, which will
default to `FALSE` (i.e. the customer-level report is produced).

To run the attribution report, use the `attribution_fit()` function, for
example:

```{r}
fractional_attribution <- attribution_fit(example_path_summary,
                                          example_path_customer_map)
                                          
# Inspect some customers
fractional_attribution %>% 
  filter(customer_id %in% c('cid_1644', 'cid_1683', 'cid_1755'))
```

See that for the customer with **cid_1683**, who had the **B > A** path to
conversion, fractribution has attributed **55.3%** of the conversion to
**channel B** and **44.7%** to **channel A**.

#### Path level only

If you just want to fractional values at the path- (not customer-) level,
set `path_level_only = TRUE`:

```{r, warning=FALSE}
path_fracs <- attribution_fit(example_path_summary,
                              path_level_only = TRUE)

# Inspect some paths
path_fracs %>% 
  filter(path %in% c('B', 'A', 'B > C', 'B > D', 'B > D > G', 'D > C > A')) %>% 
  select(path:g)
```

Notice how the value of **D** in the path **B > D** (**59.5%**) reduces
significantly when a **G** is added to the path (i.e. **B > D > G**). That is,
**D**'s value drops to **26.2%**, whereas **B**'s value remains very stable
(**40.5%** to **39.3%**).

#### Path transform method

Hidden in the two examples above was the default
`path_transform_method = 'exposure'` argument. Raw paths to conversion can be
messy, especially if your lookback period is long or you have many different
channels defined. Thus it is effective to transform paths before conducting the
counterfactual search during the attribution fit. We have five options for this:
**unique**, **exposure**, **first**, **recency**, and **frequency**, each with
various benefits specific to the use case. See `?attribution_fit` for details on
each.

To demonstrate how the `path_transform_method` will change the attribution fit,
here's the same input data from above, but using the **first** path transform
instead (which could be more appropriate for a brand awareness type marketing
strategy):

```{r}
# 'first' path transform
first_fit <- attribution_fit(example_path_summary,
                             example_path_customer_map,
                             path_transform_method = 'first')
                                          
# Inspect some customers
first_fit %>% 
  filter(customer_id %in% c('cid_1644', 'cid_1683', 'cid_1755'))
```

Notice how the attribution fractions have changed slightly. This change could be
more dramatic in your own data.

### Reporting

Now that we can get the attribution fractions at the customer-level, we
will typically want to join in revenue information and roll up to a higher-level
channel report.

Fractribution supports this with the `channel_attribution_report()` and
`channel_revenue_attribution_report()` functions, as well as a
`last_touch_attribution()` function for comparison purposes. We'll explore these
below.

#### Channel Attribution

`channel_attribution_report()` will take the path-level attribution fractions
fit in `attribution_fit()` and roll up to a channel level report. For example if
we use our `fractional_attribution` fit from before, we get the following:

```{r}
fractribution <- channel_attribution_report(fractional_attribution)

# Inspect
fractribution
```

So here we can say channel **b** was attributed **683** conversions, etc.

It's also a good sanity check to confirm that the fractribution
`attribution_conversions` and `path_summary$total_conversions` column sums are
equal:

```{r}
# Attributed conversions
sum(fractribution$attributed_conversions)

# Original paths to conversion
sum(example_path_summary$total_conversions)
```


#### Combining Last touch attribution

For comparison we include a `last_touch_attribution()` function. We can
aggregate last-touch attribution with the following: 

```{r}
last_touch <- last_touch_attribution(example_path_summary)

# Inspect
last_touch
```

Note this also includes the same `path_transform_method` argument which you can
tweak depending on your use case. It defaults to **unique** by which we get the
actual last touch.

We can join this to the fractribution channel fractions using the following
(note there is a little processing needed on channel first):

```{r}
# Clean up last-touch channel name
library(stringr)
last_touch <- last_touch %>% 
    mutate(last_channel = str_to_lower(last_channel) %>%
      str_replace_all(' - ', ' ') %>%
      str_replace_all('-', ' ') %>%
      str_replace_all(' ', '_'))

# Join to fractribution
channel_report <- fractribution %>% 
  inner_join(last_touch, by = c('channel' = 'last_channel')) %>% 
  mutate(difference = attributed_conversions - last_touch_conversions)

# Inspect
channel_report
```

You can now see which channels are shown to improve in value in fractribution
over last-touch attribution. For example **28** additional conversions were
granted to **b** (last-touch undervaluing). Also **17** conversions were removed
from **c** (last-touch overvaluing), etc.

#### Adding a default conversion value

Next, you might have some default conversion value you can assign to
estimate an attributed revenue:

```{r}
# Set default conversion value
conversion_val <- 70

# Distribute 
channel_report_default_rev <- channel_report %>% 
  mutate(attributed_revenue = attributed_conversions * conversion_val,
         last_touch_revenue = last_touch_conversions * conversion_val) %>% 
  select(attributed_conversions, attributed_revenue,
         last_touch_conversions, last_touch_revenue)
  
# Inspect
channel_report_default_rev
```

**Note** we don't recommend this approach but _instead_ suggest using the actual
revenue if possible (see the next section).

#### Channel report with revenue and ROAS

Attributing a default, single, conversion value to all customers is suboptimal
if you know—from your own transaction systems—what the actual conversion was for
each customer.

If you do know this revenue, you can get the most leverage from fractribution's
customer-level output.

The `channel_revenue_attribution_report()` function will facilitate the process
for you.

It takes two additional inputs: `conversion_revenue` and (optionally)
`channel_spend`. Including `channel_spend` means you will also get a **ROAS**
calculation in the report.

Like with `attribution_fit()` there are example datasets loaded with the
package. We'll explore these below and then show the report function.

##### Conversion revenue

The `conversion_revenue` input is a dataframe with columns `customer_id` and
`conversion_value`. The `customer_id`s are from `path_customer_map` and the
`conversion_value`s are the currency value for each customer's conversion.

For your reference there is an `example_conversion_revenue` dataframe loaded
with the package:

```{r}
example_conversion_revenue
```

##### Channel spend

The `channel_spend` input is a dataframe with columns `channel` and
`total_spend` to use for ROAS calculations. Each channel that appears in a
path should have a record, with the `total_spend` being the amount spent on the
channel during the reporting period. If the channel is non-marketing (e.g.
organic search) set the `total_spend` to `NA`.

Again there is a reference, `example_channel_spend`, loaded with the
package:

```{r}
example_channel_spend
```

##### Attributed Revenue and ROAS

Finally we can calculate attributed revenue and ROAS. To do this use the
`channel_revenue_attribution_report()` function:

```{r}
fractribution_and_roas <-
  channel_revenue_attribution_report(fractional_attribution,
                                     example_conversion_revenue,
                                     example_channel_spend)

# Inspect
fractribution_and_roas
```

### Run all path transform methods for comparison

If you want to run all path transform methods and compare them you can do
something like the following code examples.

#### Compare attribution fits

Map through each `path_transform_method` and collect `attribution_fit()`s in a
list:

```{r, eval=FALSE}
# Comparing the different methods
path_transform_methods <- c("unique", "exposure", "first", "frequency")

attribution_models <- purrr::map(
  path_transform_methods,
  ~ attribution_fit(example_path_summary,
                    example_path_customer_map,
                    .x))

names(attribution_models) <- path_transform_methods

# Inspect
attribution_models
```

#### Comparing channel reports

Continuing from above, pass the list of fits (`attribution_models`) into the
`channel_attribution_report()` function, again collecting the results in a list:

```{r, eval=FALSE}
# See the final reports
channel_reports <- purrr::map(attribution_models,
                              channel_attribution_report)

# Inspect
channel_reports
```

#### Comparing last-touch attribution reports

Similarly to above, we can also map through `last_touch_attribution()` for each
`path_transform_method`:

```{r, eval=FALSE}
path_transform_methods <- c("unique", "exposure", "first", "frequency")

last_touch_reports <- purrr::map(
  path_transform_methods,
  ~ last_touch_attribution(
    example_path_summary,
    .x))

names(last_touch_reports) <- path_transform_methods

# Inspect
last_touch_reports
```

### Additional details

If you would like to discover more about the actual fractribution algorithm,
there is an **Attribution fit - Detailed overview** vignette included in the
package which will go into more detail, as well as explore more on the
differences between the `path_transform_method`s. To launch run:

```{r, eval=FALSE}
vignette('attribution_fit_details')
```

## Bugs and features

**fractribution.model** is a work in progress and so you might find some issues.
If you do, please let me know at [dbooth@google.com](mailto:dbooth@google.com)
and I'll try fix it asap!

Also if you find you need an additional feature please reach out to me and I can
scope and, if I think it's suitable, add into a future release.
