---
title: "ChannelAttribution package review"
author: "Daniel Booth"
date: "2019-03-12"
output:
  pdf_document: default
  html_notebook: default
---

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Overview

This vignette will present at a high-level exploration of the
`ChannelAttribution` package. We will then compare ChannelAttribution to
fractribution and last touch approaches.

First let's load the required package:
```{r, message=FALSE}
library(ChannelAttribution)
library(dplyr)
library(tidyr)
```

## Package data

There is a pre-loaded dataset within the package (see
`?ChannelAttribution::Data` for details).

```{r}
data(PathData)
customer_paths <- as_tibble(Data)
```

Inspect the data:
```{r}
customer_paths %>% glimpse()
```

Let's inspect one of these, see that different conversion values mean multiple
rows for the same path:

```{r}
customer_paths %>% filter(path == 'beta > beta > eta')
```

### Channel

Let's extract the set of **channels**:
```{r}
customer_paths %>%
  mutate(split_paths = stringr::str_split(path, " > ")) %>%
  select(split_paths) %>%
  unnest(split_paths) %>%
  count(split_paths, sort = TRUE)
```

Note this is a toy dataset.

## Fit the Markov Model for attribution

There is a `markov_model()` function to do this. For the help page:

```{r, eval=FALSE}
?markov_model
```

### Order of Markov Model

A Markov model determines the probability that a user will transition from
Sequence A to Sequence B based on the steps that each user takes through a site.
The contents of these sequences are determined by the Markov order, which ranges
from 0 to 4. Here are some guidelines to **determine what Markov Order is appropriate**:

* **Order 0**: Doesn’t know where the user came from or what step the user is on, only the probability of going to any page.
* **Order 1**: Looks back zero steps. You are currently at Step A (Sequence A). The probability of going anywhere is based on being at that step.
* **Order 2**: Looks back one step. You came from Step A (Sequence A) and are currently at Step B (Sequence B). The probability of going anywhere is based on where you were and where you are.
* **Order 3**: Looks back two steps. You came from Step A > B (Sequence A) and are currently at Step C (Sequence B). The probability of going anywhere is based on where you were and where you are.
* **Order 4**: Looks back three steps. You came from Step A > B > C (Sequence A) and are currently at Step D (Sequence B). The probability of going anywhere is based on where you were and where you are.

According to the paper this package is based on (see help file) the greatest
increase in accuracy is realized moving between second and third order, in
contrast to the marginal lift moving between the third and fourth.

So let's go with **order 3**.

### Fit

To fit we apply the function as follows:

```{r}
mod <- markov_model(customer_paths, var_path = 'path', 
                    var_conv = 'total_conversions',
                    var_value = 'total_conversion_value',
                    var_null = 'total_null', out_more = TRUE,
                    order = 3, sep = ">", seed = 101)
```

### Get the results

Since we set `out_more = TRUE` this will return a list with the attribution
results, the transition matrix generated and the removal effects.

```{r}
# Inspect
str(mod)
```

#### Channel attribution report

We can extract a channel attribution report from the `mod` object.

```{r}
mod$result %>% 
  mutate(perc_contribution = round(total_conversions / sum(total_conversions), 2))
```

## Compare ChannelAttribution with Fractribution

Let's compare the channel attribution model with fractribution and last touch.

Run it all on the fractribution sample dataset:

```{r}
# Load example data from Fractribution
path_summary <- fractribution.model::example_path_summary
```

Run channel attribution:

```{r}
mod_2 <- markov_model(path_summary, var_path = 'path', 
                      var_conv = 'total_conversions',
                      var_null = 'total_non_conversions',
                      out_more = TRUE, order = 3, sep = ">",
                      seed = 101)
```

Inspect:

```{r}
# Put in variable
channel_attribution <- mod_2$result %>% 
  mutate(channel_name = as.character(channel_name),
         total_conversions = round(total_conversions))

# Inspect
channel_attribution
```

Run Fractribution:

```{r}
# Load library
library(fractribution.model)

# Fit
fractional_attribution <- attribution_fit(example_path_summary,
                                          example_path_customer_map)

# Aggregate
fractribution <- channel_attribution_report(fractional_attribution)

# Inspect
fractribution
```

And run last touch:

```{r}
# Fit
last_touch <- last_touch_attribution(example_path_summary)

# Inspect
last_touch
```

Join all together to compare:

```{r}
# Make fractribution channels upper case
fractribution <- fractribution %>% 
    mutate(channel = stringr::str_to_upper(channel))

# Join all together
combined_channel_report <- fractribution %>% 
  inner_join(channel_attribution, by = c('channel' = 'channel_name')) %>% 
  inner_join(last_touch, by = c('channel' = 'last_channel')) %>% 
  select(channel,
         fractribution = attributed_conversions,
         channel_attribution = total_conversions,
         last_touch = last_touch_conversions)

# Inspect
combined_channel_report
```

Little differences, same ranking. No harm running all and comparing.
